### 2025-04-08 correlation.qmd
20250407 - fix formula

The $r$ value (`{r} petersenlab::apa(pearson_r, decimals = 2)`) is slightly negative.
The [effect size](#sec-practicalSignificance) of the association is small—for what is considered a small, medium, and large effect size for $r$ values, see @sec-effectSizeThresholds.
The $p$-value is less than .05, so it is a statistically significant correlation.
We could report this finding as follows: There was a statistically significant, weak negative association between height and fantasy points ($r[$ `{r} pearson_df` $]=$ `{r} petersenlab::apa(pearson_r, decimals = 2)`, $p < .001$); the taller a player was, the fewer fantasy points they tended to score.

@@

Please note that the $p$-value for a correlation assumes that the observations are independent—in particular, that the residuals are not correlated.
However, the observations are not independent in the `player_stats_seasonal` dataframe used above, because the same player has multiple rows—one row corresponding to each season they played.
This non-independence violates the traditional assumptions of the significance test of a correlation.
We could address this assumption by analyzing only one season from each player.
However, for simplicity, we present results from the whole dataframe.
In @sec-mixedModels, we discuss [mixed model](#sec-mixedModels) approaches that handle repeated measures and other data that violate assumptions of non-independence that are shared by correlation and [multiple regression](#sec-multipleRegression); assumptions of [multiple regression](#sec-multipleRegression) are described in @#sec-assumptionsRegression.

>>>>

The $r$ value (`{r} petersenlab::apa(pearson_r, decimals = 2)`) is slightly negative.
The [effect size](#sec-practicalSignificance) of the association is small—for what is considered a small, medium, and large effect size for $r$ values, see @sec-effectSizeThresholds.
The $p$-value is less than .05, so it is a statistically significant correlation.
We could report this finding as follows: There was a statistically significant, weak negative association between height and fantasy points ($r[`{r} pearson_df`] = `{r} petersenlab::apa(pearson_r, decimals = 2)`$, $p < .001$); the taller a player was, the fewer fantasy points they tended to score.

@@

Please note that the $p$-value for a correlation assumes that the observations are independent—in particular, that the residuals are not correlated.
However, the observations are not independent in the `player_stats_seasonal` dataframe used above, because the same player has multiple rows—one row corresponding to each season they played.
This non-independence violates the traditional assumptions of the significance test of a correlation.
We could address this assumption by analyzing only one season from each player.
However, for simplicity, we present results from the whole dataframe.
In @sec-mixedModels, we discuss [mixed model](#sec-mixedModels) approaches that handle repeated measures and other data that violate assumptions of non-independence that are shared by correlation and [multiple regression](#sec-multipleRegression); assumptions of [multiple regression](#sec-multipleRegression) are described in @sec-assumptionsRegression.
